---
title: "Histology Pipeline"
author: "Matt Bixley"
date: "10 March 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# notes from rconf ML workshop
# https://rstudio-conf-2020.github.io/dl-keras-tf/notebooks/project1-natural-images.nb.html
library(tidyverse)
library(keras)
library(ggplot2)
library(glue)

current_dir <- "stomach_march"
Sys.setenv(current_dir = current_dir)
```

# Histology
Naive classification of histology slides.

Keras/Tensor Flow naive classification of cancer outcomes using untrimmed or cleaned histology slides.
Data sourced from National Cancer Institute - Genomic Data Commons (https://portal.gdc.cancer.gov/)

Downloading the data
GDC does not currently enable direct querying of the TCGA diagnostic images for a specific project. To generate a list of the files to download, you have to first generate a manifest of all whole-slide images in TCGA (both frozen and diagnostic), filter the frozen section images in this list, and then match the identifiers against the sample identifiers (TCGA-##-####) for the project(s) of interest.

The manifest for all TCGA whole-slide images can be generated using the GDC Legacy Archive query.

Rows containing diagnostic image files can be identified using the Linux command line

cut -d$'\t' -f 2 gdc_manifest.txt | grep -E '\.*-DX[^-]\w*.'
After matching the slide filenames against the sample IDs from the clinical data for the project(s) of interest, the relevant filenames can be used with the GDC Data Transfer Tool or the GDC API.

```{bash, makedirs, eval = F}
mkdir data/$current_dir/
mkdir data/$current_dir/test/
mkdir data/$current_dir/training/
mkdir data/$current_dir/test/alive
mkdir data/$current_dir/test/censured
mkdir data/$current_dir/training/alive
mkdir data/$current_dir/training/censured

```
download and move the manifest to the current directory eg *stomach_march* and rename to *gdc_manifest.txt*, repeat for the clinical data which comes as a .tar.gz file

prune the manifest to those samples with diagnostic slides only
```{bash, manifest, eval = F}
cut -d$'\t' -f 2 data/$current_dir/gdc_manifest.txt | grep -E '\.*-DX[^-]\w*.' > data/$current_dir/getfile.txt
#gdc_manifest_diagnostic.txt
```

```{r, eval = F}
getfile <- read_delim(file = paste0("data/",current_dir,"/getfile.txt"), delim = "\t", col_names = F)
sample_id <- read_delim(file = paste0("data/",current_dir,"/gdc_manifest.txt"), delim = "\t")

sample_id %>% filter(., filename %in% getfile$X1) %>%
  write_delim(.,paste0("data/",current_dir,"/gdc_manifest_diagnostic.txt"), delim = "\t")
```



split the data into test and training sets 75:25 split

```{r splitdata}
#read manifest
sample_id <- read_delim(file = paste0("data/",current_dir,"/gdc_manifest_diagnostic.txt"), delim = "\t")
# read clinical
clinical_id <- read_delim(file = paste0("data/",current_dir,"/clinical.project-TCGA-STAD.2020-03-10/clinical.tsv"), delim = "\t")
status_id <- clinical_id %>% select(case_id, vital_status) %>% 
  distinct(case_id, .keep_all= TRUE)

status_id <- sample_id %>% left_join(.,status_id, by = c("id" = "case_id"))

## 75% of the sample size
smp_size <- floor(0.75 * nrow(sample_id))

## set the seed to make your partition reproducible
set.seed(123)

train_ind <- sample(seq_len(nrow(status_id)), size = smp_size)

train <- status_id[train_ind, ]
test <- status_id[-train_ind, ]

train_alive <- train %>% filter(vital_status == "Alive") %>% select(case_id) %>% 
  write_delim(.,paste0("data/",current_dir,"/train_alive"), col_names = F)
train %>% filter(vital_status == "Dead") %>% select(case_id) %>% 
  write_delim(.,paste0("data/",current_dir,"/train_dead"), col_names = F)

test %>% filter(vital_status == "Alive") %>% select(case_id) %>% 
  write_delim(.,paste0("data/",current_dir,"/test_alive"), col_names = F)
test %>% filter(vital_status == "Dead") %>% select(case_id) %>% 
  write_delim(.,paste0("data/",current_dir,"/test_dead"), col_names = F)
```

```{bash, movefiles, eval = F}
for file in $(cat data/$current_dir/train_alive); do mv ~/../../media/matt/0D830EB30D830EB3/$file ~/../../media/matt/0D830EB30D830EB3/training/alive; done
for file in $(cat data/$current_dir/train_dead); do mv ~/../../media/matt/0D830EB30D830EB3/$file ~/../../media/matt/0D830EB30D830EB3/training/censured; done
for file in $(cat data/$current_dir/test_alive); do mv ~/../../media/matt/0D830EB30D830EB3/$file ~/../../media/matt/0D830EB30D830EB3/alive; done
for file in $(cat data/$current_dir/test_censured); do mv ~/../../media/matt/0D830EB30D830EB3/$file ~/../../media/matt/0D830EB30D830EB3/cesured; done
```


<!--chapter:end:HistologyPipeline.Rmd-->

